---
title: "Classroom Attentiveness Monitor: Faculty Report"
subtitle: "A Statistical Analysis and Reporting Dashboard in R (23AI502PC)"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r setup, include=FALSE}
# This code chunk runs in the background.
# It sets up our environment, loads data, and does all calculations.
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# 1. Load Libraries
library(dplyr)
library(ggplot2)

# 2. Import and Clean Data
data <- read.csv("classroom_log.csv")
data$status <- as.factor(data$status)
data$timestamp <- as.POSIXct(data$timestamp, format = "%Y-%m-%dT%H:%M:%OS")
```

## 1. Executive Summary

This report provides a data-driven analysis of student engagement for the last recorded class session. The data was collected via an automated CV system (Module 1) and analyzed using R (Module 2).

```{r, echo=FALSE}
# 3. Perform High-Level Calculations (Unit II)
class_avg_score <- mean(data$attentiveness_score)
at_risk_students <- data %>%
  group_by(roll_no) %>%
  summarise(avg_score = mean(attentiveness_score)) %>%
  filter(avg_score < 0.6) # Define "at-risk" as < 60% average

total_students <- length(unique(data$roll_no))
```

* **Total Students Monitored:** `r total_students`
* **Overall Class Average Attentiveness:** `r round(class_avg_score * 100, 1)`%
* **Students with < 60% Avg. Attention:** `r nrow(at_risk_students)`

---

## 2. Class Attentiveness Over Time (Unit V)

This line graph shows the average attentiveness of the *entire class* from the beginning to the end of the session. This helps identify moments when engagement may have dropped (e.g., long lectures, difficult topics).

```{r, echo=FALSE, fig.width=10, fig.height=5}
# 4. Create Plot 2 (Line Graph)
plot2_linegraph <- data %>%
  mutate(time_bin = cut(timestamp, breaks = "20 sec")) %>%
  group_by(time_bin) %>%
  summarise(class_avg_score = mean(attentiveness_score)) %>%
  mutate(time_bin = as.POSIXct(time_bin)) %>%
  
  ggplot(aes(x = time_bin, y = class_avg_score)) +
  geom_line(color = "#0072B2", size = 1.2) +
  geom_point(color = "#0072B2", size = 2) +
  labs(
    title = "Average Class Attentiveness Over Time",
    x = "Session Time",
    y = "Average Attentiveness Score"
  ) +
  ylim(0, 1) + # Set Y-axis from 0 to 1
  theme_minimal(base_size = 14)

print(plot2_linegraph)
```

---

## 3. Individual Student Analysis (Unit II & V)

This section provides a detailed breakdown of each student's performance.

### Individual Score Distribution (Box Plot)

The box plot below compares the score distribution for every student tracked. This is the best way to quickly identify "at-risk" individuals (those with a low median score, shown by the line in the box) and those with inconsistent attention (a very tall box).

```{r, echo=FALSE, fig.width=10, fig.height=6}
# 5. Create Plot 1 (Box Plot)
plot1_boxplot <- ggplot(data, aes(x = reorder(roll_no, attentiveness_score), y = attentiveness_score, fill = roll_no)) +
  geom_boxplot() +
  labs(
    title = "Student Attentiveness Distribution",
    subtitle = "Comparing score distributions for each student",
    x = "Student Roll Number (Ordered by Median Score)",
    y = "Attentiveness Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none") +
  coord_flip() # Flip to horizontal for easier reading

print(plot1_boxplot)
```

### Detailed Student Data Table

This table provides the raw statistics for each student, including their average score, median, standard deviation (variance), and the number of data points collected for them.

```{r, echo=FALSE}
# 6. Create Student Summary Table
student_summary <- data %>%
  group_by(roll_no) %>%
  summarise(
    avg_score = round(mean(attentiveness_score), 2),
    median_score = round(median(attentiveness_score), 2),
    std_dev = round(sd(attentiveness_score), 2),
    time_points = n()
  ) %>%
  arrange(avg_score) # Sort by average score (lowest first)

# Display the table
knitr::kable(student_summary, caption = "Summary Statistics by Student (Unit II)")
```